---
title: "appusageR Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{appusageR Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## Batch Preprocessing

`appusageR` parses exported APP Usage / Screen Time text files that already
exist. It does not collect Android data, implement questionnaire CIER detection,
scrape app stores, or infer sensitive app categories automatically.

The numeric sources of truth are `duration_ms` and millisecond timestamp
columns. Formatted duration strings are retained for audit/display only.
`package_name` is the stable app key; `app_name` is retained for human
inspection and reporting.

```{r}
library(appusageR)

files <- list.files("raw_exports", pattern = "\\.txt$", full.names = TRUE)

batch_summary <- read_appusage_batch(
  files,
  output_dir = "output_cache",
  overwrite = FALSE,
  progress = TRUE
)

batch_summary
```

## Inspect Cached Outputs

```{r}
cache_files <- batch_summary$data_file[batch_summary$status == "success"]

load(cache_files[[1]])

data
```

## Project Cache Workflow

```{r}
second_level <- write_second_level_batch(
  batch_summary,
  overwrite = TRUE,
  progress = TRUE
)

project_root <- unique(stats::na.omit(batch_summary$project_root))[[1]]

qc_summary <- write_qc_metadata_batch(
  project_root,
  progress = TRUE
)
```

Second-level caches contain `data$event`, `data$episode`, and `data$daily`.
Each successful second-level cache has a paired `proc-2.json` metadata file.
QC updates `proc-2.json` and refreshes the project-level
`analytic_summary_table_proclevel-2.csv` without loading all RDA files into one
object.

Routine QC does not create `proclevel-3` or `proclevel-tmp`. `proclevel-3` is
reserved for a future final analytic layer.

Supported second-level grains by export type are:

- line exports: `data$episode` plus line-episode-derived `data$daily`;
- meta exports by default: faithful `data$event` plus Table 1 summary-derived
  `data$daily`;
- meta exports with explicit reconstruction: opt-in `data$episode`, with
  episode-derived or combined daily rows when requested;
- day and app exports: `data$daily`.

Unavailable grains are represented by zero-row tables. Every RDA cache contains
exactly one object named `data`.

## Explicit Meta Reconstruction

`parse_meta()` is faithful to the raw export and returns
`list(summary = ..., events = ...)`; it does not reconstruct episodes.

Meta reconstruction is explicit:

```{r}
meta_episodes <- reconstruct_meta_episodes(meta_records$events)

second_level_meta <- run_second_level_appusage(
  first_level_meta,
  reconstruct_meta = TRUE,
  meta_daily_source = "both"
)
```

`meta_daily_source = "summary"` is the default. `"episodes"` and `"both"`
require `reconstruct_meta = TRUE` and keep daily provenance through
`daily_source` and duration-comparison columns.

## Activity Type and Anomaly QC

Second-level tables include `activity_type` where app labels are available.
Rows with APP Usage background/media markers are flagged as `"background"`; all
other rows are `"foreground"`. This flag does not drop rows or change duration
calculations.

`write_qc_metadata_batch()` and `run_qc_appusage(project_dir = ...)` write
standard mobile-sensing QC and anomaly diagnostics into `proc-2.json`.
Anomaly QC summarizes implausible timestamps, durations, cross-date episodes,
meta duration disagreements, and export-date-span issues. These are diagnostics
only and do not delete rows by default.

## Manual App Category Enrichment

```{r}
dict <- read_app_category_dictionary(
  "apptypedict/apptyp_dictionary_v251016.xlsx"
)

category_summary <- write_app_categories_batch(
  project_root,
  dictionary = dict,
  overwrite = TRUE,
  progress = TRUE
)
```

Category enrichment writes `Level_1_Category` and `Level_2_Category` into
second-level `proc-2.rda` files and records match diagnostics in `proc-2.json`.
It first matches `package_name` to dictionary `App_UUID`, then falls back to
exact unambiguous app-name matches.

No category is inferred from app text automatically. Unmatched or conflicting
dictionary keys remain uncoded for user review.

## Day-Level QC

```{r}
day_files <- batch_summary$data_file[batch_summary$detected_type %in% c("day", "app")]

day_records <- do.call(rbind, lapply(day_files, function(path) {
  load(path)
  data[[1]]
}))

qc <- qc_appusage_day(
  day_records,
  min_nonempty_days = 7,
  require_all_weekdays = TRUE
)

valid_day_records <- filter_valid_appusage(day_records, qc)
```

## Local Reference Data Boundary

The package examples and tests do not use `reference/all_text_data/`. That
directory is reserved for future explicitly authorized full-data stress tests.
